\chapter{Conclusion} % Main chapter title
\label{con} % For referencing the chapter elsewhere, use \ref{con} 
\subsection{Learning graphical models with approximate inference}
In the paper \cite{heinemann2012cannot} we were looking at the implication of the optimization of ML objective while the inference is BP.
As was explain before \secref{sec:Bethe_ML} it turn out that the optimized objective is the Bethe ML \eqref{eq:bethe_like} and no ML \eqref{eq:ML}.
Bethe ML is a convex function of the optimized parameter $\thetav$ but not a smooth one so optimization should be feasible.
Writing the equation for the optimality of the parameters \eqref{eq:bethe_opt} gave similar condition to the moment matching equations \eqref{eq:moment_matching}.
There is, however, one big different - were there are more than one pseudo-marginals that maximize the Bethe likelihood there is no guarantees that moment matching exist.

The importance of moment matching can be thought by loss of information.
In ML running inference on the learned parameters result in the empirical marginals hence by learning all the information was saved.
On the other hand, where there is no moment matching the empirical marginals can not be inferred from the model which reveal loss of information by the learning process.
Hence moment matching is a crucial characteristic of learning GM.
But is there such empirical marginals - marginals that for all parameters will never be the only pseudo-marginal that minimize the Bethe free energy?
Can such marginals be characterize?

The surprising answer is yes.
There are marginals inside the marginal polytope(\eqref{eq:margpoly}) that for all parameters they are not the sole minimizer of the Bethe free energy - could not be stable fix point of BP.
But even if a marginal minimize the Bethe free energy, if there is no parameter that it is the only one, this parameter is l
This encourage the definition of a new set of marginals - Bethe learn-able.
A marginal is Bethe learn-able if there exist a parameter that this marginal is the only pseudo-marginal that minimize the Bethe free energy for this parameter.
Can a bounds to this 

